### 2024-10-10 paper/paper.qmd
Starter folder added

---
title: ""
author: ""
thanks: "Code and data are available at: [GitHub repo Link]"
date: today
date-format: long
format: pdf
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(here)
library(tidyverse)
options(scipen=999)
```

```{r}
#| include: false
#| warning: false
#| message: false
radiohead  <- readRDS(here("spotify_api/data/00-raw_data/radiohead.rds"))
radiohead
```

```{r}
#| warning: false
#| message: false
#| echo:    false
radiohead <- as_tibble(radiohead)

radiohead |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

```{r}
#| warning: false
#| message: false
#| echo:    false
rbind( the_national, radiohead) |>
  select(artist_name, album_release_date, valence) |>
  mutate(album_release_date = ymd(album_release_date)) |> 
  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```